\documentclass[12pt,oneside,a4paper]{report}
%-------------------------------------------------------
% PACKAGES

\usepackage{setspace} 
\doublespacing
\usepackage{float}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[toc,page]{appendix}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[colorlinks=true,linkcolor=blue, citecolor=blue]{hyperref}

% COMANDS FOR TABLE
\usepackage{dcolumn}
\usepackage{caption}
\newcommand\fnote[1]{\captionsetup{font=scriptsize, format=hang}\caption*{#1}}


% Econometrics notations
\newcommand{\cov}{\operatorname{cov}}
\newcommand{\var}{\operatorname{var}}
\newcommand{\mX}{\bm X}
\newcommand{\vx}{\bm x}
\newcommand{\vbeta}{\bm \beta}
\newcommand{\dto}{\stackrel{d}{\longrightarrow}}

%-------------------------------------------------------
% CONFIG

\graphicspath{{./figs/}}


%-------------------------------------------------------
% PRE-TEXT CONTENT

\title{Assessing Fiscal Distress in Subnational Governments: The Case of Brazil}
\author{Francisco Alves}
\date{2017}

%-------------------------------------------------------
% TEXT BODY CONTENT

\begin{document}

\begin{titlepage}
    \centering
    {\scshape\LARGE University of York\par}
    \vspace{1cm}
    {\scshape\Large Department of Economics and Related Studies\par}
    \vspace{1.5cm}
    {\huge\bfseries Assessing Fiscal Distress in Subnational Governments: The Case of Brazil\par}
    \vspace{2cm}
    {\Large Francisco Alves\par}
    {MSc Economics and Public Policy\par}
    \vfill
    supervised by\par
    Dr.~Michal \textsc{Horvath}

    \vfill

% Bottom of the page
    {\large \textbf{September 2017}\par}
    (6,387 words)
\end{titlepage}

\begin{abstract}
Decentralization in the provision of public services has made management of contingent liabilities arising from subnational governments (SGNs) an essential component of sovereign fiscal sustainability. In Brazil, if the central government must offer a guarantee for an individual borrowing operation, then the National Treasury Secretariat (NTS) must assess that entity ``payment capacity'' before the operation is authorized. The objective of this study is to evaluate the statistical significance of the fiscal indicators used by the NTS to assess the payment capacity of subnational governments. For this purpose, following the literature on parametric Early Warning Systems (EWS), we made use of logistic regression to explain fiscal distress events in Brazilian states in the 2008-2016 period. The results obtained indicate that from the fiscal ratios used by the NTS, Current liabilities / Cash and cash equivalents has the highest explanatory power. The main conclusion is that liquidity measures appear to have superior predictive performance than solvency measures to explain fiscal distress events.

\vspace{5mm}
\noindent
\textbf{Keywords}: fiscal sustainability, fiscal distress, subnational governments, early warning system, logistic regression, logit model

\vfill
\end{abstract}

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
I would first like to thank my dissertation advisor Dr. Michal Horvath for the useful initial discussions and for providing valuable feedback on the draft of this study.

I gratefully acknowledge the funding received towards my MSc from the Chevening Scholarships, the UK government’s global scholarship programme, funded by the Foreign and Commonwealth Office (FCO) and partner organisations.

I would also like to say a heartfelt thank to my family and friends, some closer and some distant, who in one way or another encouraged me during this period.

The final and special mention goes to Samira, who has been by my side throughout this MSc, living every single
minute of it, and without whom this amazing year would definitely not have been possible.
\vfill
\end{abstract}

\tableofcontents

\input{tex/01_introduction} %  words
\input{tex/02_literature-review} %  words
\input{tex/03_methods} %  words
\input{tex/04_results} %  words
\input{tex/05_conclusion} %  words

%-------------------------------------------------------
% POST-TEXT CONTENT


\bibliographystyle{agsm}
{\footnotesize
\bibliography{../references.bib}}

% \input{tex/10_appendix}

\end{document}

